Optimal (Multiway) Spatial Joins

### 2024-05-24 msj.tex
yufei

@@

As before, let $R_1, ..., R_{k-2}, H$, and $V$ be the input sets of the H-V $k$-SJ problem. Denote by $\J_1$ the set of type-1 result tuples defined in Section~\ref{sec:hv}. In this section, we aim to compute a set $\J^*$ satisfying
\myeqn{
    \J_1
    \subseteq
    \J^*
    \subseteq
    \J(R_1, ..., R_{k-2}, H, V) \label{eqn:hv:type1:S}
}
where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ. From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.

@@

    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment --- i.e., with the smallest (resp., largest) x-coordinate --- among all the segments in $V$ that cross $r$.
}
Define $r_1'$ as the rectangle bounded by the above four segments as shown in Figure~\ref{fig:hv:type1}(a) ($r_1'$ is the shaded area). Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$. Define
\myeqn{
    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}} \nn
}
which we refer to as the {\em trimmed} version of $R_1$. Computing the ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{D}$ (the first-crossing version). By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$. It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time.

@@

In the same fashion, we obtain the trimmed version $R_i'$ of $R_i$ for all $i \in [2, n]$, spending $O(k n \log n)$ time in total.

We now solve the $(k-2)$-SJ problem on $R_1', R_2, ..., R_{k-2}$ using the algorithm $\A$ supplied. Let us represent the result of this $(k-2)$-SJ as $\J(R_1', R_2, ..., R_{k-2})$. The lemma below shows that the result is essentially what we aim to compute in Step 1.

\begin{lemma} \label{lmm:hv:type1:proj}
    A $(k-2)$-tuple $(r_1', r_2, ..., r_{k-2})$ exists in $\J(R_1', R_2, ...,$ $R_{k-2})$ if and only if the $(k-2)$-tuple $(r_1, r_2, ..., r_{k-2})$ is in $\Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$, where $r_1$ is the rectangle in $R_1$ defining $r_1'$.
\end{lemma}

The proof can be found in Appendix~\ref{app:hv}. Note that the size of $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$ is bounded by that of $\J_1(R_1, ..., R_{k-2},$ $H, V)$, which in turn is bounded by $\out$. Thus, as a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ instance $R_1', R_2, ..., R_{k-2}$  can be settled in $F_{k-2}(n, \out)$ time.

\extraspacing {\bf Step 2.} We are ready to explain how to produce $\J_1(R_1, ..., R_{k-2}, H, V)$ from $\Pi_{1:k-2}(\J_1(R_1, ...,$ $R_{k-2}, H, V))$. Our algorithm for this purpose will find a set $S$ satisfying
\myeqn{
    \J_1(R_1, ..., R_{k-2}, H, V) 
    \subseteq 
    S
    \subseteq 
    \J(R_1, ..., R_{k-2}, H, V). \label{eqn:hv:type1:S}
}
Recall that $\J(R_1, ..., R_{k-2}, H, V)$ is the result of the original H-V $k$-SJ problem, which has a size of $\out$. Only the $k$-tuples in $S$ that are of Type 1 will be reported. We will achieve the purpose in $O(n \log n + |S|) = O(n \log n + \out)$ time.

\begin{figure} 
    \begin{tabular}{cc}
            \hspace{-3mm}
            \includegraphics[height=30mm]{./artwork/four-segs} &
            \hspace{-5mm}
            \includegraphics[height=37mm]{./artwork/oracle} \\ 
            \hspace{-3mm} (a) Constructing $r_1'$ from $r_1$ & 
            \hspace{-5mm} (b) Generating the set $S$ 
    \end{tabular}
    \figcapup
    \caption{Illustrations for Section~\ref{sec:hv:type1}}
    \label{fig:hv:type1}
    \figcapdown
\end{figure}

@@

Now, take an arbitrary $(k-2)$-tuple in $(r_1, ..., r_{k-2}) \in \Pi_{1:k-2}(\J_1(R_1,$ $...,$ $R_{k-2}, H, V))$. Denote by 
\myeqn{
    B = \cap_{i=1}^{k-2} r_i. \nn
}
Note that $B$ is a rectangle itself. We say that $B$ is 
\myitems{
    \item {\em bottom-bounded} by $r_a$, if the bottom edge of $B$ is a part of the bottom boundary of $r_a$ for some $a \in [k-2]$;

    \item {\em left-bounded} by $r_b$, if the left edge of $B$ is a part of the left boundary of $r_b$ for some $b \in [k-2]$.
}
Figure~\ref{fig:hv:type1}(b) shows an example with rectangles $r_1, r_2$, and $r_3$ (here $k = 5$). Their intersection is the gray rectangle $B$, which is bottom-bounded by $r_2$ and left-bounded by $r_3$.

We aim to compute 
\myitems{
    \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$; 
    \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
}
In Figure~\ref{fig:hv:type1}(b), $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2, v_3}$. As mentioned before, $r_a = r_2$ and $r_b = r_3$. We have $\cross(r_a, B; H) = \set{h_1, h_2}$ and $\cross(r_b, B; V) = \set{v_1, v_3}$. Note that $h_3$ is not in $\cross(r_a, B; H)$ because $h_3$ does not cross $B$, and $v_2$ is not in $\cross(r_b, B; V)$ because $v_2$ does not cross $r_3$.

Appendix~\ref{app:hv} proves the next property:

\begin{lemma} \label{lmm:hv:type1:cross-non-empty}
    $\cross(r_a, B; H) \ne \emptyset$ and $\cross(r_b, B; H) \ne \emptyset$.
\end{lemma}

@@

\subsection{Putting Things Together} \label{sec:hv:analysis}

\todo{Remark on Rahul's result}

>>>>

\yf{Rahul's result here}

@@

As before, let $R_1, ..., R_{k-2}, H$, and $V$ be the input sets of the H-V $k$-SJ problem. Denote by $\J_1$ the set of type-1 result tuples defined in Section~\ref{sec:hv}. In this section, we aim to compute a set $\J^*$ satisfying
\myeqn{
    \J_1
    \subseteq
    \J^*
    \subseteq
    \J(R_1, ..., R_{k-2}, H, V) \label{eqn:hv:type1:S}
}
where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ. From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest. Remember that $\out$ is the size of $\J(R_1, ..., R_{k-2}, H, V)$.

@@

    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment --- i.e., with the smallest (resp., largest) x-coordinate --- among all the segments in $V$ that cross $r$.
}
Define $r_1'$ as the rectangle bounded by the above four segments as shown in Figure~\ref{fig:hv:type1:trim} ($r_1'$ is the shaded area). We say that $r_1'$ is the {\em trimmed rectangle} of $r_1$, and conversely, $r_1$ is the {\em full rectangle} of $r_1'$. Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$. Define
\myeqn{
    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}} \nn
}
which we refer to as the {\em trimmed} version of $R_1$. Computing the ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{D}$ (the first-crossing version). By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$. It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time.

@@

In the same fashion, we obtain the trimmed version $R_i'$ of $R_i$ for all $i \in [2, n]$, spending $O(k n \log n)$ time in total. We now solve the $(k-2)$-SJ problem on $R_1', R_2', ..., R_{k-2}'$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}). This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R_1', R_2, ..., R_{k-2})$. We prove in Appendix~\ref{app:hv}:

%The lemma below shows that the result is essentially what we aim to compute in Step 1.

\begin{lemma} \label{lmm:hv:type1:proj}
    $|\J(R_1', R'_2, ..., R'_{k-2})| \le \out$.
\end{lemma}

We reminder the reader that $\out$ is the output size of the original H-V $k$-SJ problem. As a corollary of Lemma~\ref{lmm:hv:type1:proj}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.

\yf{example}

\begin{figure}
    \includegraphics[height=30mm]{./artwork/four-segs}
    \figcapup
    \caption{Trimming}
    \label{fig:hv:type1:trim}
    \figcapdown
\end{figure}

\extraspacing {\bf Generating $\bm{\J^*}$.} Take an arbitrary $(k-2)$-tuple in $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$. The reader should recall from Section~\ref{sec:bricks} that $B_\bm{t}$ is the intersection of of $r'_1, r'_2, ..., r'_{k-2}$, $\gleft(\bm{t})$ is the $r_i$ ($i \in [k-2]$) satisfying $\xleft(r'_i) = \xleft(B_\bm{t})$, and $\gbot(\bm{t})$ is the $r_i$ ($i \in [k-2]$) satisfying $\ybot(r'_i) = \ybot(B_\bm{t})$.

@@

We now introduce:
% \myitems{
%     \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$;
%     \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
% }
\myeqn{
    \cross(\bm{t}, H) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \nn \\
    \cross(\bm{t}, V) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
}
These sets have important properties, as proved in Appendix~\ref{app:hv}:

\begin{lemma} \label{lmm:hv:type1:cross-non-empty}
    Suppose that $r_1, r_2, ..., r_{k-2}$ are the full rectangles of $r_1', r_2', ..., r_{k-2}'$, respectively. All the following statements are true:
    \myenums{
        \item $\cross(\bm{t}, H) \ne \emptyset$ and $\cross(\bm{t}, V) \ne \emptyset$.
    }
\end{lemma}

\yf{example}

@@
